---
title: "Homework 6"
author: "Brian Detweiler"
date: "March 11, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(TSA)
library(forecast)
```

# 1. Consider three seperate AR(1) models: $\phi = 0.1, \phi = 0.5$, and $\phi = 0.8$.

## (a) For each model, calculate $\rho_1$ and $\rho_7$.

$$
\begin{split}
  Y_t &= 0.1 Y_{t - 1} - e_t \\
\end{split}
$$

We must find $\rho_k = \frac{\gamma_k}{\gamma_0}$. First we need $\gamma_0$ in each case.

$$
\begin{split}
  \gamma_0 &= Cov(Y_t, Y_t) = Var(Y_t) \\
  &= Var(0.1 Y_{t - 1} - e_t) \\
  &= 0.01 Var(Y_{t - 1}) + Var(e_t) \\
  &= 0.01 \gamma_0 + \sigma_e^2 \\
  \gamma_0 - 0.01 \gamma_0 &= \sigma_e^2 \\
  0.99 \gamma_0 &= \sigma_e^2 \\
  \gamma_0 &= \frac{\sigma_e^2}{0.99} \\
\end{split}
$$

More generally, $\gamma_0 = \frac{\sigma_e^2}{1 - \phi_1^2}$.

We will expand the Yule-Walker equations,

$$
\begin{split}
  \gamma_k &= \phi_1 \gamma_{k - 1} \\
  \rho_k &= \phi_1 \rho_{k - 1} \\
\end{split}
$$

So we now have

$$
\begin{split}
  \gamma_1 &= \phi_1 \gamma_0 \\
  &= 0.1 \gamma_0 \\
  &= \frac{0.1 \sigma_e^2}{0.99} \\
  &= 0.1010101 \sigma_e^2 \\
  \rho_1 &= \frac{\gamma_1}{\gamma_0} \\
  &= \frac{0.1010101 \sigma_e^2}{\frac{\sigma_e^2}{0.99}} \\
  &= 0.1010101 (0.99) \\
  &= 0.1 \\
  &= \phi_1 \\
  \rho_2 &= \phi_1 \rho_1 \\
  \rho_2 &= 0.1 (0.1) \\
  \rho_2 &= \phi_1^2 \\
\end{split}
$$

Now we can see the pattern emerging. $\rho_k = \phi_1^k$, and hence $\rho_7 = \phi_1^7 = (0.1)^7$.

The autocorrelation function shows a decaying sequence.

```{r, fig.align="center", fig.height=4, fig.width=5, echo=FALSE}
rhos <- c(1, 0.1, (0.1)^2, (0.1)^3, (0.1)^4, (0.1)^5, (0.1)^6, (0.1)^7)
plot(0:7, rhos, main = "ACF for first 7 lags for phi_1 = 0.1")
segments(x0 = 0:7, y0 = 0, x1 = 0:7, y1 = rhos)
```

\pagebreak

Now, since we've generalized it, we can easily apply this to the other models.

$$
\begin{split}
  Y_t &= 0.5 Y_{t - 1} - e_t \\
\end{split}
$$

$$
\begin{split}
  \gamma_0 &= \frac{\sigma_e^2}{1 - \phi_1^2} \\
  &= \frac{\sigma_e^2}{1 - 0.5} \\
  &= \frac{\sigma_e^2}{0.5} \\
  \gamma_1 &= \phi_1 \gamma_0 \\
  &= 0.5 \gamma_0 \\
  &= \frac{0.5 \sigma_e^2}{0.5} \\
  &= \sigma_e^2 \\
  \rho_1 &= \frac{\gamma_1}{\gamma_0} \\
  &= \frac{\sigma_e^2}{\frac{\sigma_e^2}{0.5}} \\
  &= 0.5 \\
  &= \phi_1 \\
  \rho_2 &= \phi_1 \rho_1 \\
  \rho_2 &= 0.5 (0.5) \\
  \rho_2 &= \phi_1^2 \\
  \rho_7 &= \phi_1^7 = 0.5^7 \\
\end{split}
$$

```{r, fig.align="center", fig.height=4, fig.width=5, echo=FALSE}
rhos <- c(1, 0.5, (0.5)^2, (0.5)^3, (0.5)^4, (0.5)^5, (0.5)^6, (0.5)^7)
plot(0:7, rhos, main = "ACF for first 7 lags for phi_1 = 0.5")
segments(x0 = 0:7, y0 = 0, x1 = 0:7, y1 = rhos)
```

\pagebreak

$$
\begin{split}
  Y_t &= 0.8 Y_{t - 1} - e_t \\
\end{split}
$$

$$
\begin{split}
  \rho_1 &= \phi_1 = 0.8 \\
  \rho_7 &= \phi_1^7 = 0.2097152 \\
\end{split}
$$

```{r, fig.align="center", fig.height=4, fig.width=5, echo=FALSE}
rhos <- c(1, 0.8, (0.8)^2, (0.8)^3, (0.8)^4, (0.8)^5, (0.8)^6, (0.8)^7)
plot(0:7, rhos, main = "ACF for first 7 lags for phi_1 = 0.8", ylim=c(0, 1))
segments(x0 = 0:7, y0 = 0, x1 = 0:7, y1 = rhos)
```

\pagebreak

## (b) For each model, calculate $Var(r_1)$ and $Var(r_7)$.

We define the sample ACF as 
$$
\begin{split}
  r_k &= \frac{\sum_{t = k + 1}^n (Y_t - \overline{Y}) (Y_{t - k} - \overline{Y})}{\sum_{t = 1}^n (Y_t - \overline{Y})^2} \\
\end{split}
$$

For an AR(1) process with $\rho_k = \phi^k$ for $k > 0$, 

$$
\begin{split}
  Var(r_k) &\approx \frac{1}{n} \bigg[ \frac{(1 + \phi^2)(1 - \phi^{2k})}{1 - \phi^2} - 2k\phi^{2k} \bigg] \\
\end{split}
$$

We'll create a helper function to calculate the numerical parts of the variance in \textsf{R}:

```{r}
partial.var.rk <- function(phi, k) {
  return(((1 + phi^2) * (1 - phi^(2 * k)) / (1 - phi^2) - 2 * k * phi^(2 * k)))
}
```

So for the given models and $k = 1, 7$, we have

## Model 1, $\phi_1 = 0.1$

```{r, include=FALSE}
phi <- 0.1
```

$$
\begin{split}
  Var(r_1) &\approx \frac{1}{n} \bigg[ \frac{(1 + (`r phi`)^2)(1 - (`r phi`)^{2})}{1 - (`r phi`)^2} - 2(`r phi`)^{2} \bigg] \\
  &\approx \frac{`r partial.var.rk(phi, 1)`}{n}  \\
  Var(r_7) &\approx \frac{1}{n} \bigg[ \frac{(1 + (`r phi`)^2)(1 - (`r phi`)^{14})}{1 - (`r phi`)^2} - 14(`r phi`)^{14} \bigg] \\
  &\approx \frac{`r partial.var.rk(phi, 7)`}{n}  \\
\end{split}
$$

## Model 2, $\phi_1 = 0.5$
```{r, include=FALSE}
phi <- 0.5
```

$$
\begin{split}
  Var(r_1) &\approx \frac{1}{n} \bigg[ \frac{(1 + (`r phi`)^2)(1 - (`r phi`)^{2})}{1 - (`r phi`)^2} - 2(`r phi`)^{2} \bigg] \\
  &\approx \frac{`r partial.var.rk(phi, 1)`}{n}  \\
  Var(r_7) &\approx \frac{1}{n} \bigg[ \frac{(1 + (`r phi`)^2)(1 - (`r phi`)^{14})}{1 - (`r phi`)^2} - 14(`r phi`)^{14} \bigg] \\
  &\approx \frac{`r partial.var.rk(phi, 7)`}{n}  \\
\end{split}
$$

## Model 3, $\phi_1 = 0.8$
```{r, include=FALSE}
phi <- 0.8
```

$$
\begin{split}
  Var(r_1) &\approx \frac{1}{n} \bigg[ \frac{(1 + (`r phi`)^2)(1 - (`r phi`)^{2})}{1 - (`r phi`)^2} - 2(`r phi`)^{2} \bigg] \\
  &\approx \frac{`r partial.var.rk(phi, 1)`}{n}  \\
  Var(r_7) &\approx \frac{1}{n} \bigg[ \frac{(1 + (`r phi`)^2)(1 - (`r phi`)^{14})}{1 - (`r phi`)^2} - 14(`r phi`)^{14} \bigg] \\
  &\approx \frac{`r partial.var.rk(phi, 7)`}{n}  \\
\end{split}
$$

## (c) For each model, use the \texttt{arima.sim} function to simulate a time series of length $n = 60$. Then use the \texttt{acf} function to calculate $r_1$ and $r_7$. Remember to set up a random seed for your simulation.

```{r}
set.seed(0)

phi1 <- 0.1

model1.r1 <- c()
model1.r7 <- c()
for (i in 1:999) {
  sim <- arima.sim(n = 60, model = list(ar=(phi1)))
  r <- acf(sim, plot = FALSE)
  model1.r1 <- c(model1.r1, r[[1]][1])
  model1.r7 <- c(model1.r7, r[[1]][7])
}

r <- acf(sim)
model1.r1 <- c(model1.r1, r[[1]][1])
model1.r7 <- c(model1.r7, r[[1]][7])
```

```{r}
phi1 <- 0.5

model2.r1 <- c()
model2.r7 <- c()
for (i in 1:999) {
  sim <- arima.sim(n = 60, model = list(ar=(phi1)))
  r <- acf(sim, plot = FALSE)
  model2.r1 <- c(model2.r1, r[[1]][1])
  model2.r7 <- c(model2.r7, r[[1]][7])
}

r <- acf(sim)
model2.r1 <- c(model2.r1, r[[1]][1])
model2.r7 <- c(model2.r7, r[[1]][7])
```

```{r}
phi1 <- 0.8

model3.r1 <- c()
model3.r7 <- c()
for (i in 1:999) {
  sim <- arima.sim(n = 60, model = list(ar=(phi1)))
  r <- acf(sim, plot = FALSE)
  model3.r1 <- c(model3.r1, r[[1]][1])
  model3.r7 <- c(model3.r7, r[[1]][7])
}

r <- acf(sim)
model3.r1 <- c(model3.r1, r[[1]][1])
model3.r7 <- c(model3.r7, r[[1]][7])
```

## (d) Based on your results in parts (a) and (b), are $r_1$ and $r_7$ from part (c) within 2 standard deviations of $\rho_1$ and $\rho_7$ respectively?

### $\phi = 0.1$
For an AR(1), we know that $c_{ii} = \frac{1}{n} \bigg[ \frac{(1 + \phi^2)(1 - \phi^{2i})}{1 - \phi^2} - 2i \phi^{2i} \bigg]$. Given the standard error of $r_k = \sqrt{\frac{c_{ii}}{n}}$. We've calculated these previously in part **(b)**, so we just need values for $n$.

We can investigate increasing values of $n$ to to get an idea of how the variance behaves with more data. We'll use $n = 1, 100, 1000$.

## Model 1, $\phi_1 = 0.1$

```{r, include=FALSE}
phi <- 0.1
n <- 1
```

$$
\begin{split}
  n &= 1\\
  SE(r_1) &\approx `r sqrt(partial.var.rk(phi, 1) / n)` \\
  SE(r_7) &\approx `r sqrt(partial.var.rk(phi, 7) / n)` \\
\end{split}
$$

```{r, include=FALSE}
n <- 100
```

$$
\begin{split}
  n &= 100\\
  SE(r_1) &\approx `r sqrt(partial.var.rk(phi, 1) / n)` \\
  SE(r_7) &\approx `r sqrt(partial.var.rk(phi, 7) / n)` \\
\end{split}
$$

```{r, include=FALSE}
n <- 1000
```

$$
\begin{split}
  n &= 1000\\
  SE(r_1) &\approx `r sqrt(partial.var.rk(phi, 1) / n)` \\
  SE(r_7) &\approx `r sqrt(partial.var.rk(phi, 7) / n)` \\
\end{split}
$$

## Model 2, $\phi_1 = 0.5$

```{r, include=FALSE}
phi <- 0.5
n <- 1
```

$$
\begin{split}
  n &= 1\\
  SE(r_1) &\approx `r sqrt(partial.var.rk(phi, 1) / n)` \\
  SE(r_7) &\approx `r sqrt(partial.var.rk(phi, 7) / n)` \\
\end{split}
$$

```{r, include=FALSE}
n <- 100
```

$$
\begin{split}
  n &= 100\\
  SE(r_1) &\approx `r sqrt(partial.var.rk(phi, 1) / n)` \\
  SE(r_7) &\approx `r sqrt(partial.var.rk(phi, 7) / n)` \\
\end{split}
$$

```{r, include=FALSE}
n <- 1000
```

$$
\begin{split}
  n &= 1000\\
  SE(r_1) &\approx `r sqrt(partial.var.rk(phi, 1) / n)` \\
  SE(r_7) &\approx `r sqrt(partial.var.rk(phi, 7) / n)` \\
\end{split}
$$

## Model 3, $\phi_1 = 0.8$
```{r, include=FALSE}
phi <- 0.8
n <- 1
```

$$
\begin{split}
  n &= 1\\
  SE(r_1) &\approx `r sqrt(partial.var.rk(phi, 1) / n)` \\
  SE(r_7) &\approx `r sqrt(partial.var.rk(phi, 7) / n)` \\
\end{split}
$$

```{r, include=FALSE}
n <- 100
```

$$
\begin{split}
  n &= 100\\
  SE(r_1) &\approx `r sqrt(partial.var.rk(phi, 1) / n)` \\
  SE(r_7) &\approx `r sqrt(partial.var.rk(phi, 7) / n)` \\
\end{split}
$$

```{r, include=FALSE}
n <- 1000
```

$$
\begin{split}
  n &= 1000\\
  SE(r_1) &\approx `r sqrt(partial.var.rk(phi, 1) / n)` \\
  SE(r_7) &\approx `r sqrt(partial.var.rk(phi, 7) / n)` \\
\end{split}
$$

In all cases, the standard error falls within 2 standard deviations.

## (e) Repeat part (c) for 1000 times. Draw histograms for $r_1$'s and $r_7$'s for each model. What proportion of $r_1$'s and $r_7$'s are within 2 standard deviations of $\rho_1$ and $\rho_7$?

## Model 1, $\phi_1 = 0.1$

```{r, fig.height=4, fig.width=5, fig.align='center'}
phi1 <- 0.1
lower <- phi1 - (2 * sd(model1.r1))
upper <- phi1 + (2 * sd(model1.r1))

within.sd <- (1 - sum((model1.r1 < lower) + (model1.r1 > upper)) / length(model1.r1)) * 100
hist(model1.r1, breaks=100, freq=F, xlim=c(min(lower, model1.r1), max(upper, model1.r1)))
abline(v = lower, col="red")
abline(v = upper, col="red")
```

For $r_1$ in model 1, `r within.sd`% fall within 2SE.

```{r, fig.height=4, fig.width=5, fig.align='center'}
phi1 <- (0.1)^7
lower <- phi1 - (2 * sd(model1.r7))
upper <- phi1 + (2 * sd(model1.r7))

within.sd <- (1 - sum((model1.r7 < lower) + (model1.r7 > upper)) / length(model1.r7)) * 100
hist(model1.r7, breaks=100, freq=F, xlim=c(min(lower, model1.r7), max(upper, model1.r7)))
abline(v = lower, col="red")
abline(v = upper, col="red")
```

For $r_7$ in model 1, `r within.sd`% fall within 2SE.

## Model 2, $\phi_1 = 0.5$

```{r, fig.height=4, fig.width=5, fig.align='center'}
phi1 <- 0.5
lower <- phi1 - (2 * sd(model2.r1))
upper <- phi1 + (2 * sd(model2.r1))

within.sd <- (1 - sum((model2.r1 < lower) + (model2.r1 > upper)) / length(model2.r1)) * 100
hist(model2.r1, breaks=100, freq=F, xlim=c(min(lower, model2.r1), max(upper, model2.r1)))
abline(v = lower, col="red")
abline(v = upper, col="red")
```

For $r_1$ in model 2, `r within.sd`% fall within 2SE.

```{r, fig.height=4, fig.width=5, fig.align='center'}
phi1 <- (0.5)^7
lower <- phi1 - (2 * sd(model2.r7))
upper <- phi1 + (2 * sd(model2.r7))

within.sd <- (1 - sum((model2.r7 < lower) + (model2.r7 > upper)) / length(model2.r7)) * 100
hist(model2.r7, breaks=100, freq=F, xlim=c(min(lower, model2.r7), max(upper, model2.r7)))
abline(v = lower, col="red")
abline(v = upper, col="red")
```

For $r_7$ in model 2, `r within.sd`% fall within 2SE.


## Model 3, $\phi_1 = 0.8$

```{r, fig.height=4, fig.width=5, fig.align='center'}
phi1 <- 0.8
lower <- phi1 - (2 * sd(model3.r1))
upper <- phi1 + (2 * sd(model3.r1))

within.sd <- (1 - sum((model3.r1 < lower) + (model3.r1 > upper)) / length(model3.r1)) * 100
hist(model3.r1, breaks=100, freq=F, xlim=c(min(lower, model3.r1), max(upper, model3.r1)))
abline(v = lower, col="red")
abline(v = upper, col="red")
```

For $r_1$ in model 3, `r within.sd`% fall within 2SE.

```{r, fig.height=4, fig.width=5, fig.align='center'}
phi1 <- (0.8)^7
lower <- phi1 - (2 * sd(model3.r7))
upper <- phi1 + (2 * sd(model3.r7))

within.sd <- (1 - sum((model3.r7 < lower) + (model3.r7 > upper)) / length(model3.r7)) * 100
hist(model3.r7, breaks=100, freq=F, xlim=c(min(lower, model3.r7), max(upper, model3.r7)))
abline(v = lower, col="red")
abline(v = upper, col="red")
```

For $r_7$ in model 3, `r within.sd`% fall within 2SE.

\begin{flushright}
  $\blacksquare$
\end{flushright}

\pagebreak


# 2. Consider an $AR(1)$ model with $\phi = 0.6$.

## (a) Use the \texttt{arima.sim} function to simulate three time series of lengths $n = 15$, 75, and 100.

```{r}
phi1 <- 0.6

sim1 <- arima.sim(n = 15, model = list(ar=(phi1)))
sim2 <- arima.sim(n = 75, model = list(ar=(phi1)))
sim3 <- arima.sim(n = 100, model = list(ar=(phi1)))

sim1.r <- acf(sim1)
sim1.r1 <- sim1.r[[1]][1]
sim2.r <- acf(sim2)
sim2.r1 <- sim2.r[[1]][1]
sim3.r <- acf(sim3)
sim3.r1 <- sim3.r[[1]][1]
```

## (b) For each set of simulated data, calculate $r_1$.

### $n = 15$
```{r}
sim1.r1
```

### $n = 75$
```{r}
sim2.r1
```

## $n = 100$
```{r}
sim3.r1
```
## (c) For each $n$, what is $Var(r_1)$? Is $r_1$ within 2 standard deviations of $\rho_1$ for each sample?

Using the function we defined earlier, \texttt{partial.var.rk}, we just need to substitute in our values for $n$.

### $n = 15$

$$
\begin{split}
  Var(r_1) &= `r partial.var.rk(0.6, 1) / 15` \\
\end{split}
$$

## $n = 75$

$$
\begin{split}
  Var(r_1) &= `r partial.var.rk(0.6, 1) / 75` \\
\end{split}
$$

### $n = 100$

$$
\begin{split}
  Var(r_1) &= `r partial.var.rk(0.6, 1) / 100` \\
\end{split}
$$

## (d) Repeat part (a) for 1000 times. For each $n$, draw a histogram of the 1000 $r_1$'s, and find what proportion of $r_1$'s are within 2 standard deviations of $\rho_1$.

```{r}
phi1 <- 0.6

model4.r1 <- c()
for (i in 1:999) {
  sim <- arima.sim(n = 60, model = list(ar=(phi1)))
  r <- acf(sim, plot = FALSE)
  model4.r1 <- c(model4.r1, r[[1]][1])
}

r <- acf(sim)
model4.r1 <- c(model4.r1, r[[1]][1])

lower <- 0.6 - (2 * sd(model4.r1))
upper <- 0.6 + (2 * sd(model4.r1))

within.sd <- (1 - sum((model4.r1 < lower) + (model4.r1 > upper)) / length(model4.r1)) * 100

hist(model4.r1, breaks=100, freq=F, xlim=c(min(lower, model4.r1), max(upper, model4.r1)))
abline(v = lower, col="red")
abline(v = upper, col="red")
```

`r within.sd`% of the observations fall within 2SE.

\begin{flushright}
  $\blacksquare$
\end{flushright}

\pagebreak

# 3. Consider an MA(1) model with $\theta = 0.6$

## (a) Use the \texttt{arima.sim} function to simulate three time series of lengths $n$ = 15, 75, and 150. Note that \textsf{R} uses the negative of the MA coefficients.

```{r}
theta1 <- -0.6

sim1 <- arima.sim(n = 15, model = list(ma=(theta1)))
sim2 <- arima.sim(n = 75, model = list(ma=(theta1)))
sim3 <- arima.sim(n = 100, model = list(ma=(theta1)))

sim1.r <- acf(sim1)
sim2.r <- acf(sim2)
sim3.r <- acf(sim3)
```

## (b) For each set of simulated data, calculate $r_1$.

### $n = 15$

```{r}
sim1.r[[1]][1]
```

### $n = 75$

```{r}
sim2.r[[1]][1]
```

### $n = 100$

```{r}
sim3.r[[1]][1]
```

## (c) For each $n$, what is $Var(r_1)$? Is $r_1$ within 2 standard deviations of $\rho_1$ for each sample?

For an MA(1) model, $c_{11} = 1 - 3 \rho_1^2 + 4 \rho_1^4$. 

Thus, $Var(r_1) = \frac{c_{11}}{n}$.

```{r}
rho1 <- -0.6 / (1 + (0.6)^2)
```

For an MA(1) model, $\rho_1 = \frac{- \theta}{1 + \theta^2}$. In this case, $\rho_1 = \frac{-0.6}{1 + (0.6)^2} = `r rho1`$.

$$
\begin{split}
  Var(r_1) &= \frac{c_{11}}{n} = \frac{1 - 3 \rho_1^2 + 4 \rho_1^4}{n} \\
  &= \frac{1 - 3 (`r rho1`)^2 + 4 `r rho1`^4}{n} \\
  &= \frac{`r 1 - 3 * rho1^2 + 4 * rho1^4`}{n} \\
\end{split}
$$

### $n = 15$

$$
\begin{split}
  Var(r_1) &= \frac{`r 1 - 3 * rho1^2 + 4 * rho1^4`}{15} \\
  &= `r (1 - 3 * rho1^2 + 4 * rho1^4) / 15` \\
\end{split}
$$

### $n = 75$

$$
\begin{split}
  Var(r_1) &= \frac{`r 1 - 3 * rho1^2 + 4 * rho1^4`}{75} \\
  &= `r (1 - 3 * rho1^2 + 4 * rho1^4) / 75` \\
\end{split}
$$

### $n = 100$

$$
\begin{split}
  Var(r_1) &= \frac{`r 1 - 3 * rho1^2 + 4 * rho1^4`}{100} \\
  &= `r (1 - 3 * rho1^2 + 4 * rho1^4) / 100` \\
\end{split}
$$

## (d) Repeat part (a) for 1000 times. For each $n$, draw a histogram of the 1000 $r_1$'s, and find what proportion of $r_1$'s are within 2 standard deviations of $\rho_1$.

```{r}
theta1 <- -0.6

rho1 <- -0.6 / (1 + (0.6)^2)

model5.r1 <- c()
for (i in 1:999) {
  sim <- arima.sim(n = 60, model = list(ma=(theta1)))
  r <- acf(sim, plot = FALSE)
  model5.r1 <- c(model5.r1, r[[1]][1])
}

r <- acf(sim)
model5.r1 <- c(model5.r1, r[[1]][1])

lower <- rho1 - (2 * sd(model5.r1))
upper <- rho1 + (2 * sd(model5.r1))

within.sd <- (1 - sum((model5.r1 < lower) + (model5.r1 > upper)) / length(model5.r1)) * 100

hist(model5.r1, breaks=100, freq=F, xlim=c(min(lower, model5.r1), max(upper, model5.r1)))
abline(v = lower, col="red")
abline(v = upper, col="red")
```

`r within.sd`% of the observations fall within 2SE.


\begin{flushright}
  $\blacksquare$
\end{flushright}

\pagebreak

# 4. The dataset \texttt{days} contains accounting data. The data is the number of days it took to receive payment for 130 consecutive orders from a particular distributor.

```{r}
data(days)
```

## (a) Plot the times series. Are there any unusual values?

```{r}
plot(days)
day1 <- which(days == 55)
day2 <- which(days == 49)
day3 <- which(days == 63)
```

There are three highly unusual days, at days `r day1`, `r day2`, and `r day3`.

## (b) Draw the sample ACF and sample PACF plots. What do you find?

```{r}
acf(days)
pacf(days)
```

## (c) Replace the unusual values with a value of 35 days. Redraw the sample ACF and sample PACF plots. Are they different from part (b)?

```{r}
days.tmp <- days
days.tmp[which(days.tmp == 55)] <- 35
days.tmp[which(days.tmp == 49)] <- 35
days.tmp[which(days.tmp == 63)] <- 35
plot(days.tmp)
abline(reg=lm(days.tmp~time(days.tmp)))
```

The time series appears to have a slightly decreasing trend, and although the variance appears to be constant, the mean is not, so it is not yet stationary.

## (d) What ARMA model would you specify for this series after removing the outliers? Explain.

We'll take the first difference to get a stationary time series.

```{r}
plot(diff(days.tmp))
acf.val <- acf(diff(days.tmp), lag.max = 60)
phi1 <- acf.val[[1]][1]
pacf.val <- pacf(diff(days.tmp), lag.max = 60)
theta1 <- pacf.val[[1]][1]
```

The ACF cuts off after 1 lag, so the simplest model we could choose would be an MA(1). This is confirmed using \texttt{auto.arima}.


```{r}
suggested.model <- auto.arima(days.tmp)
suggested.model
theta1 <- suggested.model$coef[[1]]
```

```{r}
p <- 0
d <- 1
q <- 1

sim1 <- arima.sim(n = 130, 
                  start.innov=(40/ -theta1), 
                  n.start=1, 
                  innov=c(0, rnorm(130 - 1)), 
                  model = list(order=c(p, d, q), 
                               sd=sd(days.tmp),
                               ma=c(-theta1)))
plot(sim1)
lines(days.tmp, col="red")
```

This model does not fit very well though. 

\begin{flushright}
  $\blacksquare$
\end{flushright}

\pagebreak