---
title: "Homework 9"
author: "Brian Detweiler"
date: "April 24, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(TSA)
library(MASS)
library(forecast)
library(ggplot2)
library(zoo)
library(scales)
```

```{r, fig.height=4, fig.width=5, fig.align='center'}
set.seed(48548493)
```

# 1. Consider an AR(1) model with $\phi$ = -0.5 and $\mu$ = 14.  If our last observation occurred at time 50 and the value observed was 12, calculate $\hat{Y}_{50}(1)$, $\hat{Y}_{50}(2)$, and $\hat{Y}_{50}(10)$.

From equation 9.3.6, we get
$$
\begin{split}
  \hat{Y}_{t}(1) &= \mu + \phi(Y_t - \mu) \\
  &= 14 - 0.5 (12 - 14) \\
  &= `r 14 - 0.5 * (12 - 14)` \\
\end{split}
$$

Once we have $\hat{Y}_{t}(1)$, we can use equation 9.3.7 in conjunction with this and obtain the general values for arbitrary $\hat{Y}_{t}(l)$.

$$
\begin{split}
  \hat{Y}_{t}(2) &= \mu + \phi(\hat{Y}_{2 - 1} - \mu) \\
  &= 14 - 0.5 (15 - 14) \\
  &= `r 14 - 0.5 * (15 - 14)` \\
\end{split}
$$

$$
\begin{split}
  \hat{Y}_{t}(10) &= \mu + \phi(\hat{Y}_{10 - 1} - \mu) \\
  &= 14 - 0.5 (14.0039062 - 14) \\
  &= `r 14 - 0.5 * (14.0039062 - 14)` \\
\end{split}
$$

\begin{flushright}
  $\blacksquare$
\end{flushright}

\pagebreak

# 2. Consider the AR(2) model $Y_t = 5 + 1.1 Y_{t - 1} - 0.5 Y_{t - 2} + e_t$ with $\sigma_e^2 = 2$.

## (a) If the last 3 observed values of the sequence are $Y_{90} = 9$, $Y_{91} = 11$, and $Y_{92} = 10$, find the forecasts $\hat{Y}_{92}(1)$, $\hat{Y}_{92}(2)$, and $\hat{Y}_{92}(3)$.

For an ARMA(2, 0), we can use equation 9.3.28, 

$$
\begin{split}
  \hat{Y}_t(l) &= \phi_1 \hat{Y}_t(l - 1) + \phi_2 \hat{Y}_{t}(l - 2) \\
  \hat{Y}_{92}(1) &= \phi_1 \hat{Y}_{92} + \phi_2 \hat{Y}_{91} \\
  &= 1.1 (10) - 0.5 (11) \\
  &= `r 1.1 * (10) - 0.5 * (11)` \\
  \hat{Y}_{92}(2) &= \phi_1 \hat{Y}_{93} + \phi_2 \hat{Y}_{92} \\
  &= 1.1 (5.5) - 0.5 (10) \\
  &= `r 1.1 * (5.5) - 0.5 * (10)` \\
  \hat{Y}_{92}(3) &= \phi_1 \hat{Y}_{94} + \phi_2 \hat{Y}_{93} \\
  &= 1.1 (1.05) - 0.5 (5.5) \\
  &= `r 1.1 * (1.05) - 0.5 * (5.5)` \\
\end{split}
$$

## (b) Find the variances of $e_{92}(1)$, $e_{92}(2)$, $e_{92}(3)$. 

$$
\begin{split}
 %e_t(1) &= Y_{t + 1} - \hat{Y}_{t}(1) = e_{t + 1}\\
  Var(e_t(l)) &= \sigma_{e}^2 \sum_{j = 0}^{l - 1} \psi_{j}^2 = l \sigma_e^2 \\
  Var(e_t(1)) &= 1 (2) = 2 \\
  Var(e_t(2)) &= 2 (2) = 4 \\
  Var(e_t(3)) &= 3 (2) = 6 \\
\end{split}
$$


## (c) Find the 95% prediction limits for the forecasts $\hat{Y}_{92}(1)$, $\hat{Y}_{92}(2)$, and $\hat{Y}_{92}(3)$.

The 95% prediction limits for the forcasts are at $\pm 2 \sqrt{\sigma_e^2}$ for each value of $Var(e_t(l))$.

$$
\begin{split}
  \hat{Y}_{92}(1) &: (`r (1.1 * (10) - 0.5 * (11)) - 2`, `r 1.1 * (10) - 0.5 * (11) + 2`) \\
  \hat{Y}_{92}(2) &: (`r (1.1 * (5.5) - 0.5 * (10)) - 2 * 2`, `r (1.1 * (5.5) - 0.5 * (10)) + 2 * 2`) \\
  \hat{Y}_{92}(3) &: (`r (1.1 * (1.05) - 0.5 * (5.5)) - 2 * 3`, `r (1.1 * (1.05) - 0.5 * (5.5)) + 2 * 3`)\\
\end{split}
$$

\begin{flushright}
  $\blacksquare$
\end{flushright}

\pagebreak

# 3. Use \texttt{arima.sim} with $n = 100$ to simulate an ARIMA(0,2,2) with $\theta_1 = 1$ and $\theta_2 = -0.75$ (Use your NUID in \texttt{set.seed}). Store the data as \texttt{y}. Look at the data, notice that there are 102 observations and the first two are both zero. Remove the first two (zero) observations.
```{r}
y <- arima.sim(n=100, list(order=c(0, 2, 2), ma = c(1, -0.75)))
plot(y)
y <- ts(y[3:102])
```

## (a) Fit an ARIMA(0,2,2) model to \texttt{y[1:95]}, the first 95 observations of the simulated series, and find the maximum likelihood estimates of $\theta_1$ and $\theta_2$ .

```{r}
y.95 <- ts(y[1:95])
fit <- arima(y.95, order=c(0, 2, 2), method = "ML")
fit
```

## (b) Construct a time series plot that shows observations 91 through 95, and the forecasts (with prediction limits) for observations 96 through 100.

```{r}
fore <- forecast(y.95, model = fit, h = 5)
plot(fore)
```

## (c) Use \texttt{points(96:100, y[96:100], col="red")} to add the actual observations to the plot. Compare the forecasts with the actual observations.

```{r}
plot(fore)
points(96:100, y[96:100], col="red")
```

\begin{flushright}
  $\blacksquare$
\end{flushright}

\pagebreak

# 4. We have previously fitted an AR(3) to the square-root of the \texttt{hare} data.

```{r}
data(hare)
```

## (a) Fit the model and plot the time series along with the forecasts (with prediction limits) for the next 12 observations of the square root of hare abundance.

```{r}
fit <- Arima(y = sqrt(hare), order=c(3, 0, 0))
fore <- forecast(object=sqrt(hare), model=fit, h=12)
plot(fore)
```
## (b) Use your answer to part (a) to find the numerical values of the forecasts for the hare abundance for the next 12 years.

```{r}
hares <- (coredata(fore$mean))^2
upper <- (coredata(fore$upper[,2]))^2
lower <- (coredata(fore$lower[,2]))^2

for (i in 1:12) {
  print(paste0("Year: ", (i + 1935), " - ", hares[i], " hares"))
}
```

## (c) How do we plot the values you found in (b) along with the corresponding intervals? The plot command has a \texttt{transform} option that allows the data to be transformed. 

```{r}
square <- function(x) {
  y <- x^2
}

sq <- square(fore$fitted)

plot(sq, xlim=c(1905, 1948))
lines(1936:1947, hares, col="red")
lines(1936:1947, upper, col="blue")
lines(1936:1947, lower, col="blue")
```

\begin{flushright}
  $\blacksquare$
\end{flushright}

\pagebreak


# 5. Consider the multiplicative seasonal ARIMA $(0, 0, 2) \times (0, 1, 0)_{4}$ model.

## (a). Write down the model $Y_t$ = _____________________.

$$
\begin{split}
  W_{t} &= \nabla_4 Y_{t} = Y_{t - 4} + e_{t} - \theta_1 e_{t - 1} - \theta_2 e_{t - 2}) \\
\end{split}
$$


## (b). Find the first four $\psi$-weights for this model.

$$
\begin{split}
  Y_{t} &= Y_{t - 4} + e_t - \theta_1 e_{t - 1} - \theta_2 e_{t - 2} \\
  &= (Y_{t - 8} + e_{t - 4} - \theta_1 e_{t - 5} - \theta_2 e_{t - 6}) + e_{t} - \theta_1 e_{t - 1} - \theta_2 e_{t - 2} \\
  &= e_t - \theta_1 e_{t - 1} - \theta_2e_{t - 2} + e_{t - 4} - \theta_1 e_{t - 5} - \theta_2 e_{t - 6} + Y_{t - 8} \\
\end{split}
$$

Thus the $\psi$-weights here are $\psi_1 = 1$, $\psi_2 = -\theta_1$, $\psi_3 = 0$, and $\psi_4 = 1$.

## (c). Suppose that $\theta_1 = 0.5$, $\theta_2 = -0.25$, and $\sigma_e^2 = 1$, and that the last four observed values were: 25, 20, 25, 40, with corresponding residuals: 2, 1, 2, 3. Predict the next 4 values.

We have
$$
\begin{split}
  \hat{Y}_t(1) &= Y_{t - 3} - \theta_1 e_t - \theta_2 e_{t - 1} \\
  &= 25 - (0.5)(3) - (-0.25) (2) \\
  &= 24 \\
  \hat{Y}_t(2) &= Y_{t - 2} - \theta_2 e_t \\
  &= 20 - (- 0.25)(3) \\
  &= 20.75 \\
  \hat{Y}_t(3) &= Y_{t - 1} = 25 \\
  \hat{Y}_t(4) &= 40 \\
\end{split}
$$

## (d) Construct prediction intervals for the predictions found in (b).

$$
\begin{split}
  \hat{Y}_t(1) &: 24 \pm 2 \sqrt{1} \Rightarrow (20, 26) \\
  \hat{Y}_t(2) &: 20.75 \pm 2 \sqrt{1 + (0.5)^2} \Rightarrow (`r 20.75 - (2 * sqrt(1 + (0.5)^2))`, `r 20.75 + (2 * sqrt(1 + (0.5)^2))`) \\
  \hat{Y}_t(3) &: 24 \pm 2 \sqrt{1 + (0.5)^2 + (0.25)^2} \Rightarrow  (`r 24 - (2 * sqrt(1 + (0.5)^2 + (0.25)^2))`, `r 24 + (2 * sqrt(1 + (0.5)^2 + (0.25)^2))`) \\
  \hat{Y}_t(4) &: 40 \pm 2 \sqrt{1 + (0.5)^2 + (0.25)^2 + 0^2} \Rightarrow  (`r 40 - (2 * sqrt(1 + (0.5)^2 + (0.25)^2))`, `r 40 + (2 * sqrt(1 + (0.5)^2 + (0.25)^2))`) \\
\end{split}
$$

\begin{flushright}
  $\blacksquare$
\end{flushright}

\pagebreak

# 6. The dataset \texttt{JJ} contains the earnings per share for each quarter from 1960 to 1980 for Johnson and Johnson.

```{r}
data(JJ)
```

## (a) Plot the time series and also the logarithm of the series. Whether should we use the log transformation to model this series? Explain.

```{r}
plot(JJ)
plot(log(JJ))
```

We'll use the log, because it creates a linear pattern.

## (b) Based on your decision of part (a), find the most appropriate ARIMA $(p, d, q) \times (P, D, Q)_s$ to fit the data or transformed data.

```{r}
fit <- auto.arima(log(JJ))
```

We will use an ARIMA $(2, 0, 0) \times (1, 1, 0)_4$.

## (c) Estimate the parameters of the chosen model.

```{r}
fit
```

## (d) Use your model to predict the next 4 values of the series.

```{r}
fore <- forecast(object = JJ, model=fit, h = 4)
plot(fore)
fore
```


\begin{flushright}
  $\blacksquare$
\end{flushright}

\pagebreak
