---
title: "Homework 5"
author: "Brian Detweiler"
date: "March 2 , 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(TSA)
```

# 1. For each of the following, identify it as an ARIMA model. That is, find the values of $p, d$, and $q$ and the values of the parameters ($\phi$'s and $\theta$'s). Recall that by definition ARMA($p, q$) models must be stationary and invertible.

## (a) $Y_t = 0.6 Y_{t - 1} + 0.4 Y_{t - 2} + e_t - 0.5 e_{t - 1} + 0.25 e_{t - 2}$

This appears to be an ARMA(2, 2), with $\phi_1 = 0.6$ and $\phi_2 = 0.4$, $\theta_1 = -0.5$ and $\theta_2 = 0.25$. 

We must verify the assumptions that it is stationary and invertible.

We have

$$
\begin{split}
  \phi_1 + \phi_2 &= 0.6 + 0.4 = 1.0 \nless 1.0\\
  \phi_2 - \phi_1 &= 0.4 - 0.6 = -0.2 < 1 \\
  |\phi_2| &= 0.4 < 1 \\
\end{split}
$$

Here the first constraint is violated, so 

## (b) $Y_t = 2 Y_{t - 1} - Y_{t - 2} + e_t$

## (c) $Y_t = 0.5 Y_{t - 1} - 0.5 Y_{t - 2} + e_t - 0.1 e_{t - 1}$

\begin{flushright}
  $\blacksquare$
\end{flushright}

\pagebreak


# 2. For each ARIMA model described in Question 1, find the numerical values of $\psi_0, \psi_1, \psi_2, \psi_3, \psi_4$ and a recurrence relation for $\psi_k, k > 4$.


\begin{flushright}
  $\blacksquare$
\end{flushright}

\pagebreak

# 3. Consider a stationary process $\{Y_t\}$. Show that if $\rho_1  < 0.5$ then $\nabla Y_t$ has a larger variance than $Y_t$.


\begin{flushright}
  $\blacksquare$
\end{flushright}

\pagebreak

# 4. The data set \texttt{gold} from the \texttt{TSA} library contains the daily price of gold for 252 trading days in 2005.

```{r}
data(gold)
```

## (a) Construct a time series plot of the price of gold $Y_t$. What are the interesting features of this process?

```{r, fig.height=4, fig.width=5, fig.align='center'}
plot(gold)
```

The price of gold does not seem to be based on a deterministic trend, as we can see it begins to increase in variance after 150 days.

## (b) Let $W_t = \nabla (\ln Y_t)$, the differences of the logarithms. Construct a time series plot of $W_t$. Does it look stationary?

```{r, fig.height=4, fig.width=5, fig.align='center'}
plot(diff(log(gold)))
abline(h=-0.02)
abline(h=0.02)
```

This looks heteroskedastic, and therefore not stationary. However, the variance is actually quite small.

## (c) Use the sample ACF to investigate whether $W_t$ is a white noise process.

```{r, fig.height=4, fig.width=5, fig.align='center'}
acf(diff(log(gold)))
```

## (d) Investigate whether $W_t$ is a normal white noise process.


\begin{flushright}
  $\blacksquare$
\end{flushright}

\pagebreak
