
  
#ggplot(diff(dat.grouped), aes(y=VIOLATIONS, x=VIOLATION.DATE)) +
#geom_line() +
#geom_hline(yintercept = violations.mean, col="blue") +
#geom_hline(yintercept = violations.mean + two.sd, col="red", linetype="dashed") +
#geom_hline(yintercept = violations.mean - two.sd, col="red", linetype="dashed") +
#theme(legend.position="none") +
#labs(title="All Traffic Violations in Chicago (outliers removed)", x="Violation Date", y="Number of Violations")

We still have a problem with day-to-day variance. This can be seen by aggregating to the day of week.

```{r, fig.height=3, fig.width=3, fig.align='center', echo=FALSE}
ZOO <- zoo::zoo(dat.grouped)
zoo.df <- as.data.frame(ZOO)
zoo.df$VIOLATIONS <- as.numeric(as.character(zoo.df$VIOLATIONS))
zoo.df$VIOLATION.DATE <- as.Date(zoo.df$VIOLATION.DATE)
zoo.df$DOW <- lubridate::wday(zoo.df$VIOLATION.DATE, label=T)
violations.dow <- summarise(dplyr::group_by(zoo.df, DOW), sum(VIOLATIONS)) 
colnames(violations.dow) <- c('DOW', 'COUNT')

#ggplot(violations.dow, aes(x=DOW, y=COUNT)) +
  #geom_point() +
  #geom_line(aes(group = 1)) +
  #labs(title="Violations by day of week", x="Day of week", y="Total Violations")
```

Here it can be seen that violations increase throughout the business week, peaking on Friday, when people are rushing to get home to start their weekend. This would suggests a seasonal trend, however, after further investigation, attempting to fit a model to the series as it is results in a very poorly fit, largely due to the day-to-day variance.

One way we can get around this is to model each day of the week as its own time series. 

```{r, fig.height=3, fig.width=3, fig.align='center', echo=FALSE, warning=FALSE}
m.df <- melt(zoo.df, id.vars=c("VIOLATION.DATE", "VIOLATIONS"))
head(m.df)

sunday.df <- zoo.df %>% filter(DOW == 'Sun')
monday.df <- zoo.df %>% filter(DOW == 'Mon')
tuesday.df <- zoo.df %>% filter(DOW == 'Tues')
wednesday.df <- zoo.df %>% filter(DOW == 'Wed')
thursday.df <- zoo.df %>% filter(DOW == 'Thurs')
friday.df <- zoo.df %>% filter(DOW == 'Fri')
saturday.df <- zoo.df %>% filter(DOW == 'Sat')

ggplot() +
  geom_line(data = sunday.df, aes(x=VIOLATION.DATE, y=VIOLATIONS, color="Sunday")) +
  geom_line(data = monday.df, aes(x=VIOLATION.DATE, y=VIOLATIONS, color="Monday")) 


#+
#  geom_line(data = tuesday.df, aes(x=VIOLATION.DATE, y=VIOLATIONS, color="Tuesday")) +
#  geom_line(data = wednesday.df, aes(x=VIOLATION.DATE, y=VIOLATIONS, color="Wednesday")) +
#  geom_line(data = thursday.df, aes(x=VIOLATION.DATE, y=VIOLATIONS, color="Thursday")) +
#  geom_line(data = friday.df, aes(x=VIOLATION.DATE, y=VIOLATIONS, color="Friday")) +
#  geom_line(data = saturday.df, aes(x=VIOLATION.DATE, y=VIOLATIONS, color="Saturday")) 
#  
zoo.df.m <- melt(zoo.df, id.vars=c("VIOLATION.DATE", "VIOLATIONS"))

plot(sunday.df$VIOLATION.DATE, sunday.df$VIOLATIONS, type="l", col="black", ylim=c(0, 6000))
lines(monday.df$VIOLATION.DATE, monday.df$VIOLATIONS, type="l", col="blue")
lines(tuesday.df$VIOLATION.DATE, tuesday.df$VIOLATIONS, type="l", col="green")
lines(wednesday.df$VIOLATION.DATE, wednesday.df$VIOLATIONS, type="l", col="purple")
lines(thursday.df$VIOLATION.DATE, thursday.df$VIOLATIONS, type="l", col="brown")
lines(friday.df$VIOLATION.DATE, friday.df$VIOLATIONS, type="l", col="red")
lines(saturday.df$VIOLATION.DATE, saturday.df$VIOLATIONS, type="l", col="orange")



fit <- auto.arima(xts(sunday.df$VIOLATIONS, order.by = sunday.df$VIOLATION.DATE))
fit
fit.aic <- fit$aic
model <- fitted(fit)

plot(fit$x, col="red")
lines(fitted(fit), col="blue")

resid.mean <- mean(fit$residuals)
resid.2sd <- 2 * sd(fit$residuals)
plot(fit$residuals, main="ARIMA(1, 1, 3) residuals")
abline(h=resid.mean, col="red")
abline(h=resid.2sd, col="blue")
abline(h=-resid.2sd, col="blue")

lm.fit <- lm(fit$x ~ fit$residuals)
summary(lm.fit)
plot(fit$residuals, fit$x)
abline(lm.fit)
qqn <- qqnorm(residuals(fit))
qqline(residuals(fit))
r.squared <- cor(qqn$x, qqn$y)
shapiro.test(fit$residuals)
hist(fit$residuals, breaks=100)









sunday.df$VIOLATIONS <- log(sunday.df$VIOLATIONS)
sun.log <- log(sunday.df$VIOLATIONS)
mon.log <- log(monday.df$VIOLATIONS)
tue.log <- log(tuesday.df$VIOLATIONS)
wed.log <- log(wednesday.df$VIOLATIONS)
thu.log <- log(thursday.df$VIOLATIONS)
fri.log <- log(friday.df$VIOLATIONS)
sat.log <- log(saturday.df$VIOLATIONS)

sunday.xts <- xts(x = sun.log, order.by = sunday.df$VIOLATION.DATE)

fit <- auto.arima(sunday.xts)
fit
fit.aic <- fit$aic
fit.aic
model <- fitted(fit)

plot(fit$x, col="red")
lines(fitted(fit), col="blue")

resid.mean <- mean(fit$residuals)
resid.2sd <- 2 * sd(fit$residuals)
plot(fit$residuals, main="ARIMA(1, 1, 3) residuals")
abline(h=resid.mean, col="red")
abline(h=resid.2sd, col="blue")
abline(h=-resid.2sd, col="blue")


lm.fit <- lm(fit$residuals ~ fit$x)
summary(lm.fit)

plot(as.vector(sunday.xts[,1]), fit$residuals)
abline(lm.fit)
qqn <- qqnorm(residuals(fit))
qqline(residuals(fit))
r.squared <- cor(qqn$x, qqn$y)
shapiro.test(fit$residuals)
hist(fit$residuals, breaks=100)


```


```{r, fig.height=4, fig.width=5, fig.align='center', echo=FALSE}
#violations.xts <- xts(zoo.df$VIOLATIONS, zoo.df$VIOLATION.DATE)
#violations.xts.delta <- diff(log(violations.xts))
#violations.xts.delta.delta <- diff(diff(log(violations.xts)))

#lambda <- BoxCox.lambda(violations.xts) # = 0.27
#violations.xts.boxcox <- BoxCox(violations.xts, lambda)

#acf(violations.xts.boxcox, na.action = na.pass)

#pacf(violations.xts.boxcox, na.action = na.pass)

```

Here, we first attempt to transform the data to remove some of the variance. We use a Box-Cox transformation with $\lambda =$ `r lambda`.

We attempt to fit an ARIMA(1, 1, 3) model to the data and then check the fit.

```{r, fig.height=4, fig.width=5, fig.align='center', echo=FALSE}
fit <- auto.arima(violations.xts.boxcox)
fit
fit.aic <- fit$aic
model <- fitted(fit)

plot(fit$x, col="red")
lines(fitted(fit), col="blue")

resid.mean <- mean(fit$residuals)
resid.2sd <- 2 * sd(fit$residuals)
plot(fit$residuals, main="ARIMA(1, 1, 3) residuals")
abline(h=resid.mean, col="red")
abline(h=resid.2sd, col="blue")
abline(h=-resid.2sd, col="blue")

lm.fit <- lm(fit$x ~ fit$residuals)
summary(lm.fit)
plot(fit$residuals, fit$x)
abline(lm.fit)
qqn <- qqnorm(residuals(fit))
qqline(residuals(fit))
r.squared <- cor(qqn$x, qqn$y)
shapiro.test(fit$residuals)
hist(fit$residuals, breaks=100)

#autoplot(violations.xts) +
  #geom_point(data = model) +
  #labs(title="First Difference for All Violations")


# autoplot(violations.xts) +
  # geom_point(data = model) +
  # labs(title="First Difference for All Violations")


  #fit <- auto.arima(violations.ts)

#autoplot(diff(violations.ts)) +
  #labs(title="First difference for All Violations")

#lines(fit)

#acf(residuals(fit))


# decompose(violations.ts, frequency=7, start=1)
# lag.plot(dDat$resid, lags=12)

```

The residuals fit a Q-Q Plot with an R-squared of `r r.squared`. The residuals are highly normal. The AIC, however is very high. The variance on the original model is still very high, and is affecting the goodness-of-fit. We will need to deal with this.


# Camera-level Analysis

## TODO: Once the aggregated model has been fit, the same process can be applied to the individual cameras

To get a feel for the data, we'll look at a single camera. Camera 149 has the highest violation rate, so we'll look at this one.

```{r, fig.height=4, fig.width=5, fig.align='center', echo=FALSE}
#dat.1 <- dat %>% filter(CAMERA.ID == 'CHI149')
#ZOO <- zoo(dat.1$VIOLATIONS, order.by=dat.1$VIOLATION.DATE)

# zoo.df <- as.data.frame(ZOO)
# zoo.df$Date <- rownames(zoo.df)
# zoo.df$DOW <- lubridate::wday(zoo.df$Date, label=T)


# violations.dow <- summarise(dplyr::group_by(zoo.df, DOW), sum(ZOO)) 
# colnames(violations.dow) <- c('DOW', 'COUNT')

# weekday.violations <- as.data.frame(violations.dow %>% filter(DOW != 'Sat' & DOW != 'Sun') %>% select(COUNT))
# weekday.mean <- mean(weekday.violations$COUNT)
# weekday.var <- var(weekday.violations$COUNT)
# weekday.sd <- sd(weekday.violations$COUNT)


# ggplot(violations.dow, aes(x=DOW, y=COUNT)) +
  # geom_bar(stat="identity")


# violations.ts <- ts(ZOO)

# autoplot(diff(violations.ts)) +
  # labs(title="First difference for CHI149")

# acf(diff(violations.ts))

# auto.arima(violations.ts)

# str(violations.ts)

# plot(violations.ts[1:40])
# autoplot(violations.ts)
# mean(violations.ts)
# mean(diff(violations.ts))
# var(diff(violations.ts))

# d.arima <- auto.arima(violations.ts)
# d.forecast <- forecast(d.arima, level = c(95), h = 100)

# autoplot(d.forecast)
```

Looking at the ACF of the first difference of Camera 149, there appears to be a clear seasonal component. We can also run a forecast with a 100 day horizon, however due to the variance, our confidence interval will be very large the further out we go.
